PM Mungbean meta-analysis

### 2021-03-09 02_Preliminary_analysis.Rmd
Wordsmithing and appearance

```{r PA_Libraries, include=FALSE, message=FALSE}
if (!require("pacman"))
   install.packages("pacman")
pacman::p_load(tidyverse,
               lubridate,
               bomrang,
               openxlsx,
               devtools,
               reshape2,
               lme4,
               kableExtra,
               flextable,
               here,
               gridExtra)

@@

Now we have the clean data let's summarise the main variables.

@@

```
### Trial locations
The location of the trials including DMI fungicides.

```{r FinalTrialsForInclusion}
PM_MB_dat %>%
   distinct(trial_ref, location, year) %>%
   group_by(location) %>%
   summarise(n = n())%>%
   flextable()
```
<br>

@@

### Summary plots
#### Seasonal disease pressure

@@

#### Seasonal grain yield

@@

```{r}
#library(gridExtra)
# save plots into a 4 x 4 grid for publication

@@

The data plotted below are from `r length(unique(PM_MB_dat$trial_ref))` field trials between (`r min(PM_MB_dat$year)` - `r max(PM_MB_dat$year)`) of which the details are described in the [Trials considered for inclusion in meta-analysis]. 
<br>

@@

#### Yield effect
```{r TebuVPropi_Yi}
PM_MB_dat %>%
   ggplot(aes(y = grain_yield.t.ha., x = as.factor(fungicide_ai))) +
   xlab("Fungicide treatment") +
   ggtitle(label = "Grain yield for each fungicide active ingredient") +
   labs(caption = paste("Control n:", n_C,
                        "; Propiconazole n:", n_P,
                        "; Tebuconazole n:", n_T, sep = "")) +
   geom_boxplot() +
   scale_colour_usq()
```
<br>  
<br>

@@

Doses for propiconazole range from 62.5 to 125, a large variation.  
Let's inspect the difference in yields for each dose.

@@

This dose effect should be acknowledged in the meta-analysis.

@@

*****
### Mungbean cultivars

@@

Mungbean Jade shows the best quantitative resistance to powdery mildew.  
Based on a GRDC report by Sue Thompson [-@SueThompson2016] we want to know what the yield loss was in on cultivar Jade in the 2016 Trials.

@@

Some experiments were designed to investigate the effect of row spacing and plant density on powdery mildew disease as well as fungicide timing and efficacy.
The results showed that the row spacing had no statistically significant effect on powdery mildew, but narrower rows in most cases increased yield significantly.
This finding has also been shown by several other researchers' work in Queensland and New South Wales as well.
Eight trials used a row spacing of 0.75&nbsp;meters and tebuconazole as an active ingredient (AI).

@@

Let's plot the row spacing treatments for the response variables, *yield* and *disease severity*.  
The main questions are:

   * Were there any statistical differences for mungbean yield or powdery mildew severity between row spacing treatments?

@@

Let's plot disease severity in boxplots grouped by row spacing treatment.
We are asking here, which row spacing leads to the higher disease severity?

@@

```{r RS_severity_x_trial}
# Which row spacing leads to the higher disease severity
PM_MB_dat %>%
   filter(year == 2017 |
             year == 2018) %>%
   filter(fungicide_ai == "control") %>%
   ggplot(aes(y = PM_final_severity, x = factor(row_spacing))) +
   geom_boxplot(fill = usq_cols("usq charcoal"), alpha = 0.5) +
   facet_grid(cols = vars(location)) +
   ggtitle("Powdery mildew variation between different row spacing") +
   ylab("Final severity rating") +
   xlab("Row spacing (m)")
```

@@

Overall row spacing may influence both powdery mildew severity and yield.
However, in the studies that were conducted, no interaction was found between row spacing, powdery mildew severity and yield.
That is, the row spacing appeared to increase yield even if the amount of powdery mildew increased and no relationship could be detected between an interaction of the three.

@@

### Disease pressure  
The amount of disease pressure between seasons is another variable which could impact mungbean yield meta-analysis effect size. 
```{r D_pres}
# Assess the effect of yield pressure on the spray schedule efficacy
# Quantile to separate disease pressure into High, Medium and low disease pressure categories
DP33 <- 
   PM_MB_dat %>%
   filter(fungicide_ai == "control") %>%
   pull(PM_final_severity) %>%
   quantile(0.3333)

@@

>>>>

```{r PA_Libraries, echo=TRUE, results='hide'}
if (!require("pacman"))
   install.packages("pacman")
pacman::p_load(tidyverse,
               lubridate,
               bomrang,
               openxlsx,
               devtools,
               reshape2,
               lme4,
               kableExtra,
               flextable,
               here,
               gridExtra)

@@

@@

### Trial locations
The location of the trials including DMI fungicides.

```{r FinalTrialsForInclusion}
PM_MB_dat %>%
   distinct(trial_ref, location, year) %>%
   group_by(location) %>%
   summarise(n = n())%>%
   flextable()
```

@@

### Summary plots
#### Seasonal disease pressure in the control

@@

#### Seasonal grain yield in the control

@@

```{r include = FALSE}
#library(gridExtra)
# save plots into a 4 x 4 grid for publication

@@

The data plotted below are from `r length(unique(PM_MB_dat$trial_ref))` field trials between (`r min(PM_MB_dat$year)` - `r max(PM_MB_dat$year)`) of which the details are described in the [Trials considered for inclusion in meta-analysis].

@@

#### Yield effect
```{r TebuVPropi_Yi}
PM_MB_dat %>%
   ggplot(aes(y = grain_yield.t.ha., x = as.factor(fungicide_ai))) +
   xlab("Fungicide treatment") +
   ggtitle(label = "Grain yield for each fungicide active ingredient") +
   labs(caption = paste("Control n:", n_C,
                        "; Propiconazole n:", n_P,
                        "; Tebuconazole n:", n_T, sep = "")) +
   geom_boxplot() +
   scale_colour_usq()
```

<br>  
<br>  
<br>

@@

Doses for propiconazole range from 62.5&nbsp;g to 125&nbsp;g, a large variation.  
Let's inspect the difference in yields for each dose.

@@

This dose effect should be investigated in the meta-analysis.

@@

### Mungbean cultivars

@@

Cultivar Jade shows the best quantitative resistance to powdery mildew.  
Based on the 2016 GRDC report by Sue Thompson [-@SueThompson2016] we want to know what is the possible yield loss for this moderately susceptible cultivar.

@@

Some experiments were designed to investigate the effect of row spacing and plant density on powdery mildew disease as well as fungicide timing and efficacy.
The results showed that the row spacing had no statistically significant effect on powdery mildew, but narrower rows in most cases increased yield significantly.
This finding has also been shown by several other researchers' work in Queensland and New South Wales as well.

@@

Will plotting row spacing treatments show differences for the response variables, *yield* and *disease severity*.  
The main questions are:

   * Were there any statistical differences for mungbean *yield* or *powdery mildew severity* between row spacing treatments?

@@

@@

```{r RS_severity_x_trial}
# Does trial year and location interact between row spacing amd disease severity
PM_MB_dat %>%
   filter(year == 2017 |
             year == 2018) %>%
   filter(fungicide_ai == "control") %>%
   ggplot(aes(y = PM_final_severity, x = factor(row_spacing))) +
   geom_boxplot(fill = usq_cols("usq charcoal"), alpha = 0.5) +
   facet_grid(cols = vars(location)) +
   ggtitle("Powdery mildew variation between different row spacing") +
   ylab("Final severity rating") +
   xlab("Row spacing (m)")
```

@@

Overall row spacing may influence both powdery mildew severity and yield.
However, in the trial reports, no significant interaction was found between row spacing, powdery mildew severity and yield.
That is, with narrower row spacing appeared to increase yield even if the amount of powdery mildew increased and no relationship could be detected between an interaction of the three.

@@

### Disease pressure  
The amount of disease pressure between seasons is another variable which could impact mungbean yield meta-analysis effect size. 
```{r D_pres}
# Assess the effect of yield pressure on the spray schedule efficacy
# Use quantiles to separate disease pressure into High, Medium and low disease pressure categories
DP33 <- 
   PM_MB_dat %>%
   filter(fungicide_ai == "control") %>%
   pull(PM_final_severity) %>%
   quantile(0.3333)

@@

unique(PM_MB_dat[,c("location","lat","lon")])
PM_MB_dat[is.na(PM_MB_dat$lat),]

trials <- unique(PM_MB_dat, c("location","lat","lon"))
apply(trials,1, function(LL) {

   stat_sweep <- bomrang::sweep_for_stations(as.numeric(c(LL["lat"],LL["lon"])))

   # Remove banned stations
   banned_stations <- c("055286") # stations don't report rain
   stat_sweep <- stat_sweep[stat_sweep$site != banned_stations,]

   statID <- paste(stat_sweep[1,"name"],stat_sweep[1,"site"], sep = "_")

   return(statID)

})
bomrang::sweep_for_stations(unlist(PM_MB_dat[1,c("lat","lon")]))

bomrang::get
